
\chapter{Installation and Use}

\section{How to Install CGAL}
Please have a look at the {\tt README} and {\tt INSTALL} files
in the distribution.

\section{Compile Time Switches\label{CompilerFlags}}

For debugging it is reasonable to have as much internal checking as
possible.  When entering a function it may be useful to check whether
a set of preconditions is fulfilled. When leaving a function it may be 
useful to check some postconditions.  However, when your code goes into 
production mode you would like to switch all this checking 
off, as it may be expensive; Imagine an incremental convex hull algorithm
checking before and after each incremental step if the current object
is convex.

The switches are compile time switches.  By default they are switched on, 
as it is the case for the standard {\tt assert()} macro. You can switch them
off by putting a {\tt -U<switch>} in the {\tt CFLAGS} in your {\tt makefile}.

\subsubsection*{{\tt CGAL\_CHECK\_PRECONDITIONS}}
Before the execution of a function preconditions may be tested.
The preconditions are only of logical nature and never related to 
precision problems. 

\subsubsection*{{\tt CGAL\_CHECK\_POSTCONDITIONS}}
After the execution of a function postconditions may be tested.
The preconditions are only of logical nature and never related to 
precision problems.

\subsubsection*{{\tt CGAL\_CHECK\_ASSERTIONS}}
During the execution of a function assertions may be verified.
The preconditions are only of logical nature and never related to 
precision problems.

\subsubsection*{{\tt CGAL\_CHECK\_EXACTNESS\_PRECONDITIONS}}
Before the execution of a function postconditons that assume exact
computation may be tested.  Checking preconditions can be useful, but
if you use a numbertype as {\tt double}, you sooner or later run in
trouble due to inexact computations. An example: The function
\CCstyle{CGAL_Segment_2<R>::collinear_has_on(CGAL_Point_2<R> &p)} has
the precondition that the point $p$ is on the line supporting
the segment. The chance that this precondition test fails is high.

\subsubsection*{{\tt CGAL\_CHECK\_EXACTNESS\_POSTCONDITIONS}}
After the execution of a function postconditons that assume exact
computation may be tested.

\subsubsection*{{\tt CGAL\_CHECK\_DEGENERACY}}
Signals an error as soon as a degenerate object is created.

%\medskip
%There are further compile time switches that are not related to debugging.

%\subsubsection*{{\tt CGAL\_IO}}
